05_DESeq_manipulation

### 2019-12-01 analysis/05_DESeq2_manipulate.Rmd
working on cleaning up repo

@@

```{r dds, cache = T}
dds.female_hypothalamus <- subsetDESeq("female_hypothalamus")
dds.female_pituitary <- subsetDESeq("female_pituitary" )
dds.female_gonad <- subsetDESeq("female_gonad" )
dds.male_hypothalamus <- subsetDESeq("male_hypothalamus" )
dds.male_pituitary <- subsetDESeq("male_pituitary"  )
dds.male_gondad <- subsetDESeq("male_gonad")
```

## Calculate and plot total DEGs

```{r totalDEGs, eval = T}
#create list of groups for deseq contrasts
group1 <- c("m.inc.d3","m.inc.d8", "m.inc.d9", "m.inc.d17", "prolong","extend", "m.n2" ) 
group2 <- group1

a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
c <- plottotalDEGs(dds.female_gonad, "female gonad")
d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
f <- plottotalDEGs(dds.male_gondad, "male gonad")

plot_grid(a + theme(legend.position = "none"),
          b + theme(legend.position = "none"),
          c,
          d + theme(legend.position = "none"),
          e + theme(legend.position = "none"),
          f,
          nrow = 2, rel_widths = c(0.3, 0.3, 0.4)) 
```

## Calculate and plot principal components

```{r pca, eval = F}
plotPCAs(dds.female_hypothalamus, "female hypothalamus")
plotPCAs(dds.female_pituitary, "female pituitary")      
plotPCAs(dds.female_gonad, "female gonad")
plotPCAs(dds.male_hypothalamus, "male hypothalamus")
plotPCAs(dds.male_pituitary, "male pituitary")
plotPCAs(dds.male_gondad, "male gonad")
```

## heamap with minimum pvalue

```{r pheatmap, eval = F}
makepheatmap(dds.female_hypothalamus, "female hypothalamus")
makepheatmap(dds.female_pituitary, "female pituitary")
makepheatmap(dds.female_gonad, "female gonad")
makepheatmap(dds.male_hypothalamus, "male hypothalamus")
makepheatmap(dds.male_pituitary, "male pituitary")
makepheatmap(dds.male_gondad, "male gonad")        
```

## candidate genes

```{r candidates, eval = F}
plotcandidates(dds.female_hypothalamus, "female hypothalamus")
plotcandidates(dds.female_pituitary, "female pituitary")
plotcandidates(dds.female_gonad, "female gonad")
plotcandidates(dds.male_hypothalamus, "male hypothalamus")
plotcandidates(dds.male_pituitary, "male pituitary")
plotcandidates(dds.male_gondad, "male gonad")

@@

## offspring removal

```{r offspringremoval, eval = F}
g <- plotremoval(DEGs.female_hypothalamus, "Female hypothalamus", "bottom")
h <- plotremoval(DEGs.female_pituitary, "Female pituitary", "none")
i <- plotremoval(DEGs.female_gonad, "Female gonad",  "none")
j <- plotremoval(DEGs.male_hypothalamus, "Male hypothalamus", "none")
k <- plotremoval(DEGs.male_pituitary, "Male pituitary", "none")
l <- plotremoval(DEGs.male_gondad, "Male gonad", "none")

legend <- get_legend(g)

mybarplots <- plot_grid(g + theme(legend.position = "none"), h , i , j ,k , l, nrow = 2)

mybarplots2 <- plot_grid(mybarplots, legend, ncol = 1, rel_heights = c(1,0.1))
mybarplots2

pdf("../figures/sexes/offspringremoval-1.pdf", width = 10, height = 6)
plot(mybarplots2)
dev.off()

## prolong extend

```{r prolongextend, eval = F}
g <- plotprolongdelay(DEGs.female_hypothalamus, "Female hypothalamus", "bottom")
h <- plotprolongdelay(DEGs.female_pituitary, "Female pituitary", "none")
i <- plotprolongdelay(DEGs.female_gonad, "Female gonad",  "none")
j <- plotprolongdelay(DEGs.male_hypothalamus, "Male hypothalamus", "none")
k <- plotprolongdelay(DEGs.male_pituitary, "Male pituitary", "none")
l <- plotprolongdelay(DEGs.male_gondad, "Male gonad", "none")

legend <- get_legend(g)

mybarplots <- plot_grid(g + theme(legend.position = "none"), h , i , j ,k , l, nrow = 2)

mybarplots2 <- plot_grid(mybarplots, legend, ncol = 1, rel_heights = c(1,0.1))
mybarplots2

pdf("../figures/sexes/prolongextend-1.pdf", width = 10, height = 6)
plot(mybarplots2)
dev.off()

@@

a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
                         color = colData$treatment, 
                         shape = colData$tissue,
                         alpha = colData$sex)) +
  geom_point(size = 2)  + 
  mytheme() +
  theme(legend.title = element_blank(), 
        axis.text = element_blank()) +
  scale_alpha_manual(values = c(0.5,1)) +
  scale_color_manual(values = colorsmaniplevels1) +
  labs(x = "PC1: 63.6% of variance", y = "PC2: 30.1% of variance")

b <-  fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank", 
                   repel = TRUE ,
                  select.var= list(contrib = 6)) + 
      mytheme() + 
      labs(x = " ", y = " ", subtitle = NULL, title =  NULL)

png('../figures/manipulation/pcaprep-1.png',width=600,height=600, units="px",bg = "transparent")
print(a)
dev.off()

png('../figures/manipulation/pcaprep-2.png',width=600,height=600, units="px",bg = "transparent")
print(b)
dev.off()
```

>>>>

library(BiocParallel)
register(MulticoreParam(6))

@@

```{r dds, cache = T}
dds.female_hypothalamus <- subsetDESeq(m.colData, m.countData, "female_hypothalamus")
dds.female_pituitary <- subsetDESeq(m.colData, m.countData, "female_pituitary" )
dds.female_gonad <- subsetDESeq(m.colData, m.countData, "female_gonad" )
dds.male_hypothalamus <- subsetDESeq(m.colData, m.countData, "male_hypothalamus" )
dds.male_pituitary <- subsetDESeq(m.colData, m.countData, "male_pituitary"  )
dds.male_gondad <- subsetDESeq(m.colData, m.countData, "male_gonad")
```

## volcano plots

```{r}

dds.hypothalamus <- subsetDESeq2(m.colData,  m.countData, c("female_hypothalamus","male_hypothalamus"))
dds.hypothalamus <- subsetDESeq2(m.colData,  m.countData, c("female_pituitary","male_pituitary"))
dds.hypothalamus <- subsetDESeq2(m.colData,  m.countData, c("female_gonad","male_gonad"))

@@

```{r dds, cache = T}

vsdh <- vst(dds.hypothalamus, blind=FALSE)
vsdp <- vst(dds.pituitary, blind=FALSE)
vsdg <- vst(dds.gonad, blind=FALSE)

vsdfh <- vst(dds.female_hypothalamus, blind=FALSE)
vsdfp <- vst(dds.female_pituitary, blind=FALSE)
vsdfg <- vst(dds.female_gonad, blind=FALSE)
vsdmh <- vst(dds.male_hypothalamus, blind=FALSE)
vsdmp <- vst(dds.male_pituitary, blind=FALSE)
vsdmg <- vst(dds.male_gondad, blind=FALSE)
```

@@

a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
                         color = colData$treatment, 
                         shape = colData$tissue)) +
  geom_point(size = 2)  + 
  mytheme() +
  theme(legend.title = element_blank(), 
        axis.text = element_blank()) +
  scale_alpha_manual(values = c(0.5,1)) +
  scale_color_manual(values = colorsmanip) +
  labs(x = "PC1: 63.6% of variance", y = "PC2: 30.1% of variance")

b <-  fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank", 
                   repel = TRUE ,
                  select.var= list(contrib = 6)) + 
      mytheme() + 
      labs(x = " ", y = " ", subtitle = NULL, title =  NULL) +
  theme( axis.text = element_blank())

plot_grid(a,b, rel_widths = c(2,1))
```
